---
title: "The Relationship Between Educational Level and The Number of First Cousin Marriages"
---

Welcome to my project page.

Keep an eye on this space to stay updated with my project activities.

# 1. Project Overview and Scope

In this project, I am investigating the relationship between the level of education and the change in first cousin marriages over time in 81 provinces of Turkey. Specifically, I am testing the hypothesis that provinces with higher secondary and higher education completion rates exhibit lower cousin marriage rates.

# 2. Data

## 2.1 Data Source

-   [**Cousin Marriage Data**](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=W9qPaB3UHC3APCfqBcL9Pw/IbmcxfqPm7j6mUvKNbwRbeVSTs7PsxUrOw0WvM9zO)

    Exported from the Turkish Statistical Institute database, this table contains, for each province and year, the total number of marriages, the number of marriages between first cousins, and the proportion of cousin marriages (%). Coverage: 2010-2024.

-   [**Education Level Data**](https://data.tuik.gov.tr/Bulten/DownloadIstatistikselTablo?p=MOssQM1731aBCox6XZONouk1GkwqqxQ86jdpyOHxqxSLKf59B9XajpUZlq3bPMtF):

    Also from TUIK, this table provides the count of individuals by education level (illiterate, literate without diploma, primary school, secondary school, high school & equivalents, universities, master, doctorate) and sex, for each province and year. Coverage: 2010-2023.

## 2.2 General Information About Data

-   **Cousin Marriage Dataset**
    -   Columns per year: total marriages, cousin marriages, and cousin-marriage proportion (%)\
    -   Rows: 81 provinces/14 years
-   **Education Dataset**
    -   Columns: year, province code & name, total population by education level and sex\
    -   Education levels: illiterate, literate without diploma, primary, lower secondary, high school, universities, master, doctorate,unknown\
    -   Rows: 81 provinces /8 education categories with 2 sexes / 15 years

## 2.3 Reason of Choice

chose these data sets because it is necessary to examine the level of education from both a cultural and social (inbreeding) point of view. Demonstrating an inverse relationship between education levels and cousin-marriage rates may support targeted interventions in areas with low education levels. Moreover, provincial granularity offers insights into regional policy needs.

## 2.4 Preprocessing

-   **Loading & Header Processing**

    1.  Read both Excel files using `readxl::read_excel(..., col_names = FALSE)` to preserve raw header rows.

    ```{r}
    library(readxl)
    akraba_raw <- read_excel("akraba_evliligi.xlsx",
                             col_names = FALSE)

    egitim_raw <- read_excel("egitim_durumu.xlsx",
                             col_names = FALSE)
    ```

When the data obtained from TUIK are examined without manipulation, it is seen that its structure is not suitable for reading in R.

I have brought two data into tidy form by making Reshaping to make it tidy

-   **Education Data Tidy**\
    Clean headers and reshape education data into `(Province, Year, Education_level_sex,)`\
    **Cousin Marriage Data Tidy** Similarly clean headers and reshape cousin marriage data into `(Province, Year, Number of marriages, Number of marriages between first cousins, Proportion of marriages between first cousins (%))`

```{r}

tidy_education <- read_excel("tidy_education.xlsx")

tidy_marriages <- read_excel("tidy_marriages.xlsx")
```

Data recorded as R.Data and column names are shown.

```{r}

save(tidy_marriages, tidy_education, file = "data.RData")
  
# tidy_marriages
print(names(tidy_marriages))

# tidy_education
print(names(tidy_education))

```

# 3. Analysis

## 3.1 Exploratory Data Analysis

```{r Available-Years, echo=TRUE}

marriage_years  <- sort(unique(tidy_marriages$Year))
education_years <- sort(unique(tidy_education$Year))

cat("Marriages data covers years:", marriage_years, "\n")
cat("Education data covers years:", education_years, "\n")
```
```{r data-overview, echo=TRUE}

# --- Provinces ---
provinces <- sort(unique(tidy_marriages$Province))
cat("Number of provinces in marriages data:", length(provinces), "\n")
cat("Sample provinces:", paste(head(provinces, 10), collapse = ", "), "...\n\n")

# --- Education Levels  ---
edu_cols   <- names(tidy_education)[-(1:2)]
edu_levels <- unique(sub("_(Total|Male|Female)$", "", edu_cols))
cat("Education levels available:", paste(edu_levels, collapse = ", "), "\n")

```
## 3.2 Trend Analysis

xxxx

## 3.3 Model Fitting

xxxxxx

## 3.4 Results

xxxxxx

# 4. Results and Key Takeaways

xxxxxx
